AHABus \& FCORE: A Modular Bus System for High-Altitude Balloon Missions

### 2017-04-13 chapters/04-results.tex
Add radio discussion

@@

Figure \ref{fig:dit-alt} shows the altitude of the AHABus platform reported by
the received headers over the approximately two hours and twenty minutes of the
simulation. The altitude marks are matched to the time at which each packet
is received, and not emitted.

\begin{figure}[h]
\begin{tikzpicture}
\begin{axis} [
xlabel=Mission Elapsed Time (\SI{}{\second}),
ylabel=Altitude (\SI{}{\metre}),
mark size=1.0pt,
mark repeat={5},
legend style ={
    at={(0.98,0.95)}, 
    anchor=north east, draw=black, 
    fill=white,align=left
}
]
\addplot+ table [
x=MET, y=altitude, col sep=comma
] {data/test-ditl/loc.csv};
\end{axis}
\end{tikzpicture}
\centering
\caption{DITL Test - Altitude over time}
\label{fig:dit-alt}
\end{figure}

@@

Figure \ref{fig:ditl-bandwidth} shows that the average data rate used by the
platform stabilised early into the flight close to \SI{10}{bps}, or
\SI{110}{bps} (out of a maximum bandwidth of \SI{200}{bps}). Over the full
simulation's duration, an average of \SI{0.63}{Bps} were lost due to
transmission errors.

>>>>

\section{Introduction}

@@

Figure \ref{fig:ditl-alt} shows the altitude of the AHABus platform reported by
the received headers over the approximately two hours and twenty minutes of the
simulation. The altitude marks are matched to the time at which each packet
is received, and not emitted.

\begin{figure}[h]
\begin{tikzpicture}
\begin{axis} [
xlabel=Mission Elapsed Time (\SI{}{\second}),
ylabel=Altitude (\SI{}{\metre}),
mark size=1.0pt,
mark repeat={5},
legend style ={
    at={(0.98,0.95)}, 
    anchor=north east, draw=black, 
    fill=white,align=left
}
]
\addplot+ table [
x=MET, y=altitude, col sep=comma
] {data/test-ditl/loc.csv};
\end{axis}
\end{tikzpicture}
\centering
\caption{DITL Test - Altitude over time}
\label{fig:ditl-alt}
\end{figure}

@@

### 2017-04-13 chapters/05-discussion.tex
Add radio discussion

@@

@@

>>>>

% Finally, it was observed during all tests that the loss of a frame could be
% reported twice in the reception statistics: the last decoded frame number is
% only changed when a valid frame is parsed, and the frame sequence number is
% discarded if the frame contains too many erroneous bytes and cannot be
% corrected. While this avoid tracking corrupt frame sequence numbers in the
% decoder, it means that a corrupt frame will not be registered in the sequence,
% and the decoder will report a dropped frame when

@@

The payload bus

@@

The main goal of the day-in-the-life test was to evaluate whether the platform
was viable for full-duration \acrlong{hab} missions. In that respect, the test
can be considered successful: as shown in figure \ref{fig:ditl-alt}, the
simulation spanned a whole mission sequence, from launch through burst altitude,
to descent and landing, over a slightly more than two hours and twenty minutes.

While actions were taken for the test to mirror an actual mission as much as
possible, like the use of fake \acrshort{gnss} messages, one effect in
particular that had to be ignored was the weakening of the radio signal: in a
real mission, the increase in the distance between the ground station and the
platform over time means that the radio channel qualities (signal-to-noise
ratio, line of sight interruptions) are not constant. For this test, neither the
transmission nor the reception antenna were moved. Range resilience would have
to be tested through dedicated tests before the platform is used on real
missions.

The test allowed a long duration check of the radio protocol and systems. Figure
\ref{fig:ditl-bandwidth} shows that the amount of corrupted frames was well
below the total amount of data that was received. The average rate of
transmission, however, was only slightly above half of the available bandwidth,
despite the use of the calculation of the optimal intervals using the tool
discussed in section \ref{ssec:conf-file}.

This disparity between actual and optimal use of the bandwidth was caused by the payloads: the time division formula used to compute the polling interval solves for the case where each payload always transmit exactly the maximum amount of data that can be contained in packet. During the test, one of the payload generated data at a low enough rate that it fit into small packets that could be encapsulated into a single frame, reducing the required bandwidth. This issue would be challenging to address with the current architecture since it requires the polling intervals to be fixed. A better use of the bandwidth could be approached by selecting appropriate payload priority levels during the mission design phase, or designing payload to use the full capacity of the radio communications layer.

Finally, table \ref{tab:ditl-health} shows that even though the data bus
performed well overall, both test payloads were marked as unresponsive in
nearly \SI{10}{\percent} of the system health packets transmitted by the
platform. The source of the problem could not be identified due to the lack of
time and the difficulty of debugging this subsystem in particular: the
Wemos D1 development board that was used as the \acrlong{obc} does not provide
useful printing devices that could have been used to send log messages.
